---
title: "CUIP Fully Data Analysis"
author: "JD Long"
date: "8/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Docker Setup

Fire up a 64GB digital ocean docker machine: 

```
docker-machine create --digitalocean-size "s-16vcpu-64gb" --driver digitalocean --digitalocean-access-token $DO_PAT cuipchallenge
```
assumes there is an environment variable named DO_PAT that contains the DO access token.

then run

```
eval $(docker-machine env cuipchallenge) 
```
to set up some environment variables so we can connect

`docker run -d -e PASSWORD=XXXXXXXXX -p 8787:8787 rocker/tidyverse`
will start the Docker instance and use XXXXXXXXX as the Rstudio password

`docker-machine ip cuipchallenge` 
shows the IP address... now connect to that IP:8787


if a DO droplet is killed from DO, then delete the machine locally with: 
`rm -rf ~/.docker/machine/drivetimegetter`

## Fetch Data

Download the full dataset from utccuip.com:

```{r}

url <-"http://cscdc-2019.utccuip.com/CSCDC+Dataset.zip"
dest_file <- "01_data/CSCDC+Dataset.zip"

download.file(url, dest_file)
unzip(zipfile = dest_file, exdir = "01_data/")

```

